Exploratory: Optimal Resolution

### 2021-08-05 exploratory/exploratory.Rmd
color,avg roc,top10 imp

library(tidyverse)
library(knitr)
library(kableExtra)
library(ggpubr)
library(nationalparkcolors)
library(data.table)
library(cowplot)

@@

```{r data_setup}
# Variables
models <- c("rf","glmnet","xgbTree","svmRadial","rpart2")
levels <- c("phylum","class","order","family","genus","otu","asv")
levels_names <- c("Phylum","Class","Order","Family","Genus","OTU","ASV")
pal <- park_palette("Everglades",length(models))
names(pal) <- models
pal2 <- park_palette("Everglades",length(models))
names(pal2) <- c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")

@@

@@

@@

@@

```{r plot_by_method, echo=F,message=F,fig.height = 3.5, fig.width = 10, fig.align = "center"}

@@

```{r plot_by_level, echo=FALSE, message=FALSE,fig.height = 6, fig.width = 10, fig.align = "center"}
df_by_level <- full_df %>% 
  select(cv_metric_AUC,AUC,method,level) %>% 
  rename(test_AUC=AUC,cv_AUC=cv_metric_AUC) %>% 
  pivot_longer(cols=c("cv_AUC","test_AUC"),names_to="auc_type",values_to="AUC") %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")))

@@

```{r plot_all, echo=F, message=F, out.width="100%",fig.height=4}
df_all <- full_df %>% 
  select(cv_metric_AUC,AUC,method,level) %>% 
  rename(test_AUC=AUC,cv_AUC=cv_metric_AUC) %>% 
  pivot_longer(cols=c("cv_AUC","test_AUC"),names_to="auc_type",values_to="AUC") %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")))

@@

```{r dada2_all, echo=F, message=F, out.width="100%",fig.height=3}
dada2_auc_df %>% 
  rename(level=type) %>% 
  bind_rows(auc_df) %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv","DADA2"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV","DADA2")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(auc_type == "test_AUC") %>% 
  ggplot(aes(x=level,y=AUC,fill=method)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot() + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    scale_fill_manual(values=pal2,name="Model")

@@

::: {.col-md-6}
```{r rf_plot,echo=F,message=F}
m <- "Random Forest"

@@

ggplot(sub_df,aes(x=factor(level,levels=c("ASV", "OTU","Genus","Family","Order","Class", "Phylum")), y=AUC)) + 
#ggplot(sub_df,aes(x=level, y=AUC)) + 
  geom_jitter(aes(fill=method,color=method),height = 0, width = 0.2,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  scale_fill_manual(values = alpha(pal2[m],0.6)) +
  scale_color_manual(values = pal2[m]) +
  scale_y_continuous(limits=c(0.45,0.87),breaks = seq(0.5,0.8,0.1),expand=c(0,0)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.y = element_blank(),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  coord_flip() +
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  annotate("text",x=levels_names,y=rep(0.85,length(levels_names)),label=c("A","B","C","D,E","D","D","E"))
ggsave("./figures/rf_with_stats.png",width = 5,height=4,units="in")

@@

sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  mutate(y.position=c(0.85,0.95,0.9,.95)) %>%
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  #geom_boxplot(aes(fill=method)) +
  geom_jitter(aes(fill=method),height = 0, width = 0.2,alpha=0.8,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.4,1),breaks = seq(0.4,1,0.1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  #ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")

@@

sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,.95,length.out = nrow(sub_stats)))
  mutate(y.position=c(0.85,0.95,0.9,.95))

@@

sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  mutate(y.position=c(1,1,1,1,1,
                      0.9,
                      0.95,0.95,0.95,0.95,
                      0.85,
                      1.05,1.05,1.05,1.05,1.05,1.05)) %>% 
  #mutate(y.position=c(0.85,0.95,0.9,.95)) %>%
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  #geom_boxplot(aes(fill=method)) +
  geom_jitter(aes(fill=method),height = 0, width = 0.2,alpha=0.8,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.4,1.05),breaks = seq(0.4,1.05,0.1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  #ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  #stat_pvalue_manual(sub_stats,label="p_value_mod") 
  stat_pvalue_manual(sub_stats,label="sig")
```

@@

::: {.row}
::: {.col-md-6}
```{r glmnet_table,echo=FALSE}
t_table %>%
  select(level,glmnet) %>% 
  mutate(glmnet=round(glmnet,digits = 3)) %>% 
  rename("Median AUC"=glmnet,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
:::

::: {.col-md-6}
```{r glmnet_plot,echo=F,message=F}
## glmnet Model Performance
m <- "Logistic Regression"

@@

```{r rf_importance_plots,echo=F,message=F,fig.show="hold",out.width="50%"}
imp_plots[["rf-phylum-1"]]
imp_plots[["rf-phylum-2"]]

@@

```{r importance_rank, message=F,echo=F}

@@

rank_top_tax <- function(data,tax){
  sub_data <- data %>%
    filter(level==tax) %>%
    group_by(level,Taxonomy,OTU) %>% 
    summarise(mean_AUC_diff = mean(AUC_diff)) %>% 
    ungroup() %>% 
    arrange(desc(mean_AUC_diff)) %>% 
    mutate(rank=seq(1,nrow(.)))

@@

ranked_tax <- bind_rows(rank_top_tax(rf_imp_tax,"phylum"),
                        rank_top_tax(rf_imp_tax,"class"),
                        rank_top_tax(rf_imp_tax,"order"),
                        rank_top_tax(rf_imp_tax,"family"),
                        rank_top_tax(rf_imp_tax,"genus"),
                        rank_top_tax(rf_imp_tax,"otu"),
                        rank_top_tax(rf_imp_tax,"asv"))

top_10_tax <- ranked_tax %>% 
  filter(rank <= 10) %>% 
  mutate(Taxonomy=case_when(level=="otu" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            level=="asv" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            TRUE ~ Taxonomy)) %>% 
  mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
  mutate(Taxonomy=paste0(Taxonomy," (",OTU,")"))  %>% 
  select(level,Taxonomy,rank) %>% 
  pivot_wider(names_from = level,values_from = Taxonomy) 
colnames(top_10_tax) <- c("Rank","Phylum","Class","Order","Family","Genus","OTU","ASV")

ranked_tax %>% 
  filter(rank <= 10) %>% 
  mutate(Taxonomy=case_when(level=="otu" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            level=="asv" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            TRUE ~ Taxonomy)) %>% 
  mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
  mutate(Taxonomy=paste0(Taxonomy," (",OTU,")"))

kable(top_10_tax,format="html",booktabs=T,align="clllllll",
      col.names = c("Rank",levels_names)) %>% 
   kable_material(c("hover","condensed"))

@@

data <- tibble(bactname = c("Staphylococcaceae", "Moraxella", "Streptococcus", "Acinetobacter"),
               OTUname = c("OTU 1", "OTU 2", "OTU 3", "OTU 4"),
               value = c(-0.5, 0.5, 2, 3))

data %>% mutate(
  name = glue("<i>{bactname}</i> ({OTUname})"),
  name = fct_reorder(name, value))  %>%
  ggplot(aes(value, name)) + 
    geom_col(alpha = 0.5) + 
    scale_fill_identity() +
    theme(
      axis.text.y = element_markdown(),
      plot.caption = element_markdown(lineheight = 1.2))

@@

pct_bins <- NULL
level_nOTU <- NULL
for(l in levels){
  sub_data <- prevalence %>% 
    filter(level==l,kept==1)
  hist <- hist(sub_data$prevalence,breaks = seq(0,1,0.25),plot=FALSE)

@@

level_nOTU <- level_nOTU %>% 
  mutate(level_name = levels_names) 
labs <- paste0(level_nOTU$level_name," (",level_nOTU$nOTU,")")
pct_bins %>%
  mutate(level=factor(level,levels=levels,labels=labs)) %>% 
  ggplot(aes(x=prev,y=pct,fill=level)) +
    geom_bar(stat = "identity") +
    facet_wrap(~level) +
    scale_y_continuous(expand=c(0,0),limits=c(0,0.8),labels = scales::percent_format()) +
    scale_x_continuous(expand=c(0,0),breaks = c(0,.25,0.50,.75,1)) +
    xlab("Prevalence in Samples") +
    ylab("Percent of Taxa") +
    theme_bw() +
    theme(legend.position = "none",
          axis.text = element_text(size=12),
          axis.title = element_text(size=14)) +
  geom_text(aes(label = paste0((round(pct,digits=2)*100),"%")),nudge_y=0.05)
ggsave("./figures/bin_prevalence.png")
```

>>>>

library(tidyverse)
library(knitr)
library(kableExtra)
library(ggpubr)
library(nationalparkcolors)
library(data.table)
library(cowplot)
library(pROC)

@@

```{r data_setup}

# Variables
models <- c("rf","glmnet","xgbTree","svmRadial","rpart2")
levels <- c("phylum","class","order","family","genus","otu","asv")
levels_names <- c("Phylum","Class","Order","Family","Genus","OTU","ASV")
pal <- park_palette("Everglades",length(models))
names(pal) <- models
pal2 <- park_palette("Everglades",length(models))
names(pal2) <- c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")

@@

-   [Feature Importance - threshold 0.9](#importance90)

@@

-   [Prevalence VS Difference in Median Abundance](#prev-diff-abund)

-   [Prevalence by Group](#prevalence-by-group)

-   [Median Non-Zero Relative Abundance by Group](#med-relabund-group)

-   [Sensitivity at 90% Specificity](#sens-90spec)

-   [Averaged ROC curves](#avg-roc)

@@

## Color Palette

```{r color-pal,include=F}
#color palette
#colors <- c("#7E362A","#D04C44","#E6A444","#6C887A","#83BCC3","#5B6B86","#754A60")
colors <- c("#D04C44","#E7834E","#E6A444","#829E88","#83BCC3","#7989A4","#B29ABA")
names(colors) <- levels

colors2 <- colors
names(colors2) <- levels_names
pie(rep(1,length(colors2)),col=colors2,labels = names(colors2))
```

@@

```{r plot_by_method,fig.height = 3.5, fig.width = 10, fig.align = "center"}

@@

```{r plot_by_level, fig.height = 6, fig.width = 10, fig.align = "center"}
df_by_level <- full_df %>% 
  select(cv_metric_AUC,AUC,method,level) %>% 
  rename(test_AUC=AUC,cv_AUC=cv_metric_AUC) %>% 
  pivot_longer(cols=c("cv_AUC","test_AUC"),names_to="auc_type",values_to="AUC") %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")))

@@

```{r plot_all, out.width="100%",fig.height=4}
df_all <- full_df %>% 
  select(cv_metric_AUC,AUC,method,level) %>% 
  rename(test_AUC=AUC,cv_AUC=cv_metric_AUC) %>% 
  pivot_longer(cols=c("cv_AUC","test_AUC"),names_to="auc_type",values_to="AUC") %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree")))

@@

```{r dada2_all, out.width="100%",fig.height=3}
dada2_auc_df %>% 
  rename(level=type) %>% 
  bind_rows(auc_df) %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv","DADA2"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV","DADA2")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(auc_type == "test_AUC") %>% 
  ggplot(aes(x=level,y=AUC,fill=method)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot() + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    scale_fill_manual(values=pal2,name="Model")

@@

::: {.col-md-6}
```{r rf_plot}
m <- "Random Forest"

@@

ggplot(sub_df,aes(x=factor(level,levels=c("ASV", "OTU","Genus","Family","Order","Class", "Phylum")),y=AUC)) + 
  geom_jitter(aes(fill=level,color=level),height = 0, width = 0.2,shape=21,size=2) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  #scale_fill_manual(values = alpha(pal2[m],0.6)) +
  #scale_color_manual(values = pal2[m]) +
  scale_fill_manual(values = alpha(colors2,0.5)) +
  scale_color_manual(values = colors2) +
  scale_y_continuous(limits=c(0.45,0.87),breaks = seq(0.5,0.8,0.1),expand=c(0,0)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.y = element_blank(),
        axis.text = element_text(size=12),
        axis.title = element_text(size=12)) +
  coord_flip() +
  annotate("text",x=levels_names,y=rep(0.85,length(levels_names)),label=c("A","B","C","D,E","D,E","D","E")) +
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") 
ggsave("./figures/rf_with_stats.png",width = 4.5,height=5,units="in")

@@

sub_stats <- sub_stats %>% 
  mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  #mutate(y.position=c(0.85,0.95,0.9,.95)) %>%
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  #geom_boxplot(aes(fill=method)) +
  geom_jitter(aes(fill=method),height = 0, width = 0.2,alpha=0.8,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.4,1.05),breaks = seq(0.4,1,0.1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  #ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")

@@

sub_stats <- sub_stats %>% 
  mutate(y.position=seq.int(0.85,.95,length.out = nrow(sub_stats)))
  #mutate(y.position=c(0.85,0.95,0.9,.95))

@@

# sub_stats <- sub_stats %>% 
#   #mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
#   mutate(y.position=c(1,1,1,1,1,
#                       0.9,
#                       0.95,0.95,0.95,0.95,
#                       0.85,
#                       1.05,1.05,1.05,1.05,1.05,1.05)) %>% 
#   #mutate(y.position=c(0.85,0.95,0.9,.95)) %>%
#   rename(group1=level1,group2=level2)

# ggplot(sub_df,aes(x=level,y=AUC)) + 
#   geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
#   #geom_boxplot(aes(fill=method)) +
#   geom_jitter(aes(fill=method),height = 0, width = 0.2,alpha=0.8,shape=21) +
#   stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
#   scale_fill_manual(values = pal2[m]) + 
#   scale_y_continuous(limits=c(0.4,1.05),breaks = seq(0.4,1.05,0.1)) +
#   theme_bw() + xlab("") + ylab("AUROC") +
#   theme(legend.position ="none",
#         #axis.text.x = element_text(angle = 45,hjust = 1),
#         panel.grid.major.x = element_blank(),
#         axis.text = element_text(size=14)) +
#   #ggtitle(m) +
#   #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
#   #stat_pvalue_manual(sub_stats,label="p_value_mod") 
#   stat_pvalue_manual(sub_stats,label="sig")
```

@@

::: {.row}
::: {.col-md-6}
```{r glmnet_table}
t_table %>%
  select(level,glmnet) %>% 
  mutate(glmnet=round(glmnet,digits = 3)) %>% 
  rename("Median AUC"=glmnet,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
:::

::: {.col-md-6}
```{r glmnet_plot}
## glmnet Model Performance
m <- "Logistic Regression"

@@

```{r rf_importance_plots,echo=F,message=F,fig.show="hold",out.width="50%",include=FALSE}
imp_plots[["rf-phylum-1"]]
imp_plots[["rf-phylum-2"]]

@@

```{r importance_rank}

@@

rank_tax <- function(data,tax){
  sub_data <- data %>%
    filter(level==tax) %>%
    group_by(level,Taxonomy,OTU) %>% 
    summarise(mean_AUC_diff = median(AUC_diff)) %>% 
    ungroup() %>% 
    arrange(desc(mean_AUC_diff)) %>% 
    mutate(rank=seq(1,nrow(.)))

@@

ranked_tax <- bind_rows(rank_tax(rf_imp_tax,"phylum"),
                        rank_tax(rf_imp_tax,"class"),
                        rank_tax(rf_imp_tax,"order"),
                        rank_tax(rf_imp_tax,"family"),
                        rank_tax(rf_imp_tax,"genus"),
                        rank_tax(rf_imp_tax,"otu"),
                        rank_tax(rf_imp_tax,"asv"))

top_10_tax <- ranked_tax %>% 
  filter(rank <= 10)

top_10_tax_table <- ranked_tax %>% 
  filter(rank <= 10) %>% 
  mutate(Taxonomy=case_when(level=="otu" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            level=="asv" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            TRUE ~ Taxonomy)) %>% 
  mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
  mutate(Taxonomy=paste0(Taxonomy," (",OTU,")"))  %>% 
  select(level,Taxonomy,rank) %>% 
  pivot_wider(names_from = level,values_from = Taxonomy) 
colnames(top_10_tax_table) <- c("Rank","Phylum","Class","Order","Family","Genus","OTU","ASV")

# ranked_tax %>% 
#   filter(rank <= 10) %>% 
#   mutate(Taxonomy=case_when(level=="otu" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
#                             level=="asv" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
#                             TRUE ~ Taxonomy)) %>% 
#   mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
#   mutate(Taxonomy=paste0(Taxonomy," (",OTU,")"))

kable(top_10_tax_table,format="html",booktabs=T,align="clllllll",
      col.names = c("Rank",levels_names)) %>% 
   kable_material(c("hover","condensed"))

@@

# data <- tibble(bactname = c("Staphylococcaceae", "Moraxella", "Streptococcus", "Acinetobacter"),
#                OTUname = c("OTU 1", "OTU 2", "OTU 3", "OTU 4"),
#                value = c(-0.5, 0.5, 2, 3))
# ranked_tax %>% 
#   filter(rank <= 10) %>% 
#   mutate(Taxonomy=case_when(level=="otu" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
#                             level=="asv" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
#                             TRUE ~ Taxonomy)) %>% 
#   mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
#   mutate(name = glue("<i>{Taxonomy}</i> ({OTU})")) %>% 
#   ggplot(aes(mean_AUC_diff,name)) +
#     geom_col(alpha = 0.5) + 
#     scale_fill_identity() +
#     theme(
#       axis.text.y = element_markdown(),
#       plot.caption = element_markdown(lineheight = 1.2))
# 
# data %>% mutate(
#   name = glue("<i>{bactname}</i> ({OTUname})"),
#   name = fct_reorder(name, value))  %>%
#   ggplot(aes(value, name)) + 
#     geom_col(alpha = 0.5) + 
#     scale_fill_identity() +
#     theme(
#       axis.text.y = element_markdown(),
#       plot.caption = element_markdown(lineheight = 1.2))
# 
# 
# ggplot(data = diamonds) + stat_summary(
#   mapping = aes(x = depth, y = cut),
#   fun.min = function(z) { quantile(z,0.25) },
#   fun.max = function(z) { quantile(z,0.75) },
#   fun = median)

top_10_data <- inner_join(rf_imp_tax,top_10_tax,by=c("level","OTU","Taxonomy")) %>% 
  mutate(level = factor(level,levels=levels,labels=levels_names)) %>% 
  select(level,seed,OTU,Taxonomy,AUC_diff,rank) %>% 
  mutate(Taxonomy=case_when(level=="OTU" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            level=="ASV" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            TRUE ~ Taxonomy)) %>% 
  mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
  mutate(Taxonomy=gsub("_"," ",Taxonomy)) %>% 
  mutate(Taxonomy=gsub("unclassified","unclass.",Taxonomy)) %>% 
  mutate(name = glue("<i>{Taxonomy}</i> ({OTU})"))

plot_top_imp <- function(top_data,tax_level){
  top_data %>% 
    filter(level == tax_level) %>% 
    arrange(rank) %>% 
    ggplot(aes(x=AUC_diff,y=reorder(name,desc(rank)))) +
      #geom_jitter(color=colors2[tax_level],fill=alpha(colors2[tax_level],0.6)) +
      stat_summary(fun.data=median_hilow, fun.args=0.5, geom="pointrange", color=colors2[tax_level]) +
      theme_bw() +
      theme(axis.title.y = element_blank(),
        axis.text.y = element_markdown(),
        plot.caption = element_markdown(lineheight = 1.2)) +
      ggtitle(tax_level) +
      xlab("Decrease in AUC")
}

plotlist <- list()
plotlist[[1]] <- plot_top_imp(top_10_data,"Phylum")
plotlist[[2]] <- plot_top_imp(top_10_data,"Class")
plotlist[[3]] <- plot_top_imp(top_10_data,"Order")
plotlist[[4]] <- plot_top_imp(top_10_data,"Family")
plotlist[[5]] <- plot_top_imp(top_10_data,"Genus")
plotlist[[6]] <- plot_top_imp(top_10_data,"OTU")
plotlist[[7]] <- plot_top_imp(top_10_data,"ASV")

plot_grid(plotlist = plotlist,ncol = 2,align = "v")
ggsave("./figures/top10_imp.png",height = 10)

# gp1<- ggplot_gtable(ggplot_build(plotlist[[1]]))
# gp2<- ggplot_gtable(ggplot_build(plotlist[[2]]))
# gp3<- ggplot_gtable(ggplot_build(plotlist[[3]]))
# gp4<- ggplot_gtable(ggplot_build(plotlist[[4]]))
# gp5<- ggplot_gtable(ggplot_build(plotlist[[5]]))
# gp6<- ggplot_gtable(ggplot_build(plotlist[[6]]))
# gp7<- ggplot_gtable(ggplot_build(plotlist[[7]]))
# 
# maxWidth = unit.pmax(gp1$widths[2:3], 
#                      gp2$widths[2:3],
#                      gp3$widths[2:3],
#                      gp4$widths[2:3],
#                      gp5$widths[2:3],
#                      gp6$widths[2:3],
#                      gp7$widths[2:3])
# 
# gp1$widths[2:3] <- maxWidth
# gp2$widths[2:3] <- maxWidth
# gp3$widths[2:3] <- maxWidth
# gp4$widths[2:3] <- maxWidth
# gp5$widths[2:3] <- maxWidth
# gp6$widths[2:3] <- maxWidth
# gp7$widths[2:3] <- maxWidth
# 
# grid.arrange(gp1,gp2,gp3,gp4,gp5,gp6,gp7)
# ggsave("./figures/top10_imp.png")
# 
# ggarrange(plotlist = plotlist,ncol=2)

# grid.arrange(plotlist[[1]],
#              plotlist[[2]],
#              plotlist[[3]],
#              plotlist[[4]],
#              plotlist[[5]],
#              plotlist[[6]],
#              plotlist[[7]])

```
## Correlated Features
Highly correlated model features should be collapsed (i.e. permuted together) when quantifying permutation importance.
First checking if any features are correlated in this dataset.

```{r corr_feats}

get_annotated_corr <- function(tax_level){
  l_cor <- read_csv(paste0("../analysis/taxa_correlations_",tax_level,".csv"),
                    col_types = cols(.default = col_character(),
                                     corr = col_double(), 
                                     proc_corr = col_double())) 
  l_tax <- read_tsv(paste0("../data/",tax_level,"/crc.taxonomy"),
                     col_types = c(Size = col_double(),
                                   .default = col_character())) %>%
    select(-Size)

  merge <- inner_join(l_cor,l_tax,by=c("taxa1"="OTU")) %>% 
    rename(taxa1_name=Taxonomy) %>% 
    inner_join(.,l_tax,by=c("taxa2"="OTU")) %>% 
    rename(taxa2_name=Taxonomy) %>% 
    select(level,taxa1,taxa2,taxa1_name,taxa2_name,proc_corr) %>% 
    rename(corr=proc_corr)

  return(merge)
}

annotated_feature_correlations <- get_annotated_corr("phylum") %>% 
  bind_rows(get_annotated_corr("class")) %>% 
  bind_rows(get_annotated_corr("order")) %>% 
  bind_rows(get_annotated_corr("family")) %>% 
  bind_rows(get_annotated_corr("genus")) %>% 
  bind_rows(get_annotated_corr("otu")) %>% 
  bind_rows(get_annotated_corr("asv"))

maxmin_corr <- annotated_feature_correlations %>% 
  mutate(level=factor(level,levels=levels)) %>% 
  group_by(level) %>% 
  summarise(max_corr = max(corr),
            min_corr = min(corr))

corr90 <- annotated_feature_correlations %>% 
  filter(abs(corr) > 0.9)

annotated_feature_correlations %>% 
  filter(level == "genus") %>% 
  filter(abs(corr) > 0.85)

# View(annotated_feature_correlations %>% 
#        filter(level == "otu") %>%
#        filter(abs(corr) < 0.9) %>% 
#        filter(abs(corr) > 0.8))

#c("Otu00229","Otu00328","Otu00331", "Otu00361", "Otu00413", "Otu00449")
```

## Importance - threshold 0.9 {#importance90}

```{r importance90,echo=F}

@@

pct_bins <- NULL
level_nOTU <- NULL
for(l in levels){
  sub_data <- prevalence %>% 
    filter(level==l,kept==1)
  hist <- hist(sub_data$total_prevalence,breaks = seq(0,1,0.1),plot=FALSE)

@@

level_nOTU <- level_nOTU %>% 
  mutate(level_name = levels_names) 
labs <- paste0(level_nOTU$level_name," (",level_nOTU$nOTU,")")
pal <- colors2
names(pal) <- labs

pct_bins %>%
  mutate(level=factor(level,levels=levels,labels=labs)) %>% 
  ggplot(aes(x=prev,y=pct,fill=level)) +
    geom_bar(stat = "identity") +
    facet_wrap(~level,ncol=2) +
    scale_y_continuous(expand=c(0,0),limits=c(0,0.8),labels = scales::percent_format()) +
    scale_x_continuous(expand=c(0,0),breaks = seq(0,1,0.1)) +
    xlab("Prevalence in Samples") +
    ylab("Percent of Taxa") +
    theme_bw() +
    theme(legend.position = "none",
          axis.text = element_text(size=12),
          axis.title = element_text(size=14),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank()) +
  geom_text(aes(label = paste0((round(pct,digits=2)*100),"%")),nudge_y=0.08) +
  scale_fill_manual(values=pal)
ggsave("./figures/bin_prevalence.png")
```

## Prevalence VS Difference in Median Abundance {#prev-diff-abund}

Taxa prevalence across all samples compared to the difference in median relative abundance between cancer and normal samples. Top 10 taxa based on importance rank are red, rank 11-20 are blue, and the rest are grey.

```{r prev_abund,echo=F}

library(plotly)

metadata <- read_csv("../data/metadata/metadata.csv",
                     col_types = cols(.default = col_double(),
                                      sample = col_character(),
                                      Dx_Bin = col_character(),
                                      dx = col_character(),
                                      Height = col_character(),
                                      Gender = col_character(),
                                      Weight = col_character(),
                                      BMI = col_character()))

get_abund_prev <- function(tax_level,file,metadata,prev){
  sub_meta <- metadata %>%
    mutate(dx=case_when(Dx_Bin =="Adenoma" ~ "normal",
                        Dx_Bin== "Normal" ~ "normal",
                        Dx_Bin== "High Risk Normal" ~ "normal",
                        Dx_Bin== "adv Adenoma" ~ "cancer",
                        Dx_Bin== "Cancer" ~ "cancer")) %>% 
    select(sample,dx) %>% 
    drop_na() %>% 
    rename(Group=sample)
  sub_prev <- prev %>% 
    filter(level == tax_level & kept == 1) %>% 
    select(name,total_prevalence)
  abund <- as.data.frame(fread(file)) %>% 
    mutate(Group=as.character(Group)) %>% 
    select(-label,-numOtus)

  abund_prev <- inner_join(abund,sub_meta,by="Group") %>% 
    select(Group,dx,everything()) %>% 
    pivot_longer(cols=starts_with("Otu"),names_to = "name",values_to = "relabun") %>% 
    group_by(name,dx) %>% 
    summarize(median_relabund = median(relabun)) %>% 
    ungroup() %>% 
    pivot_wider(id_cols = "name",names_from="dx",values_from = "median_relabund") %>% 
    mutate(median_diff = normal - cancer) %>% 
    inner_join(.,sub_prev,by="name")

  return(abund_prev)

}

plot_abund_prev <- function(abund_prev,ranked_tax,tax_level){
  level_ranked_tax <- ranked_tax %>% 
    filter(level == tax_level) %>% 
    rename(name=OTU) %>% 
    select(-level,-mean_AUC_diff)

  ##non-interactive plot
